Notes and Problems from My Research\vspace{-1ex}

### 2019-01-18 Hopf Algebra Notes/Hopf Algebras.tex
Update Notes

\DeclareMathOperator{\Hom}{Hom}

@@

	Then using Lemma \ref{lem-func} and the fact that $A$ and $C$ are (co)algebra objects,
	we can see that the product $\ast$ satisfies the axioms required.
 \end{prf}
 \begin{lem}
	\begin{enumerate}
		\item $1_{\Hom}\ast f = m_A(u\otimes 1)(1\otimes f)(\varepsilon\otimes 1)\Delta.$
		\item $f\ast 1_{\Hom}=m_A(1\otimes u)(f\otimes 1)(1\otimes \varepsilon)$
		\item $(f\ast g)\ast h=m_A(m_A\otimes 1)((f\otimes g)\otimes h)(\Delta_C\otimes 1)\Delta_C$
		\item $f\ast(g\ast h)=m_A(1\otimes m_A)(f\otimes(g\otimes h))(1\otimes \Delta)\Delta$
	\end{enumerate}
	\label{lem-func}
 \end{lem}

@@

\end{document}

>>>>

\newcommand*{\Hom}{\ensuremath\text{Hom}}

@@

	Then using Lemma \ref{lem-func} and the fact that $A$ and $C$ are (co)algebra objects,
	we can see that the product $\ast$ satisfies the axioms required.

	Note that actually
 \end{prf}
 \begin{lem}
	\begin{enumerate}
		\item $1_{\Hom}\ast f = m_A(u\otimes 1)(1\otimes f)(\varepsilon\otimes 1)\Delta.$
		\item $f\ast 1_{\Hom}=m_A(1\otimes u)(f\otimes 1)(1\otimes \varepsilon)\Delta$
		\item $(f\ast g)\ast h=m_A(m_A\otimes 1)((f\otimes g)\otimes h)(\Delta_C\otimes 1)\Delta_C$
		\item $f\ast(g\ast h)=m_A(1\otimes m_A)(f\otimes(g\otimes h))(1\otimes \Delta)\Delta$
	\end{enumerate}
	\label{lem-func}
 \end{lem}

@@

\section{January 14, 2019}
Today we are talking about Hopf modules and later in the week we will see the fundamental
theorem of Hopf modules as well as a neat result.

Here is the ``example of the day.''
\begin{ex}
	Consider $k[x]$ with maps $\Delta(x^n)=\sum_0^nx^i\otimes x^{n-i}$ and $\varepsilon(x^n)=\delta_{1,n}$. 
	$S(x^n)=(-x)^n$. Then notice that $\Delta(x)=1\otimes x+x\otimes 1$, so $\delta(x^n)=(\delta x)^n$.

	James doesn't want to do the rest of the computations, but they can be done. :)
\end{ex}

\begin{prob}\label{prob-4.1}
	(**) Working with a Hopf algebra $V$, consider the convolution algebra $\Hom(V\otimes V,V)$
	\begin{enumerate}
		\item $(S\circ m)\ast m=1_{\Hom(V\otimes V,V)}=m\ast(m\circ s\otimes s\circ\tau)$ where $s$ is the antipode in $V$.
		\item $(\Delta\circ S)\ast \Delta=1_{\Hom(V,V\otimes V)}=\Delta\ast((s\otimes s)\circ\tau\circ \Delta)$
	\end{enumerate}
\end{prob}
\begin{sol}
	\begin{align*}
		(s\circ m)\ast m
	\end{align*}
\end{sol}
\begin{prob}
	Classify all Frobenius (to be defined) Hopf algebras of dimension 3.
\end{prob}

\subsection{Returning to (co)modules}\label{subsec-comod}
\begin{lem}
	Let $\scrT=\Vectk$, and $V$ and algebra over $k$. The following are equivalent:
	\begin{enumerate}
		\item $M$ is a left $V$-module (see last lecture).
		\item There is an action of $A$ on $M$ such that $(ab)\cdot m=a\cdot (b\cdot m)$ and $1\cdot m=m$.
		\item There is an algebra morphism $\varphi:V\to\Hom_k(M,M)$
	\end{enumerate}
\end{lem}

\begin{defn}
	Let $V$ be a coalgebra over $k$. We say $M$ is a left \textbf{comodule} if there is a map
	\[\rho_M:M\to V\otimes M\]
	such that the diagrams in figure \ref{fig-com} commute.
\end{defn}
\begin{figure}[h]\label{fig-com}
	\begin{tikzcd}
		M\ar[r,"\rho"]\ar[d,"\rho"] & V\otimes M\ar[d,"\id_V\otimes\rho"]\\
		V\otimes M\ar[r,"\Delta_V\otimes\id_M"] &V\otimes V\otimes M
	\end{tikzcd}
	\begin{tikzcd}
		M\ar[r,"\rho"]\ar[d,"\cong"] & V\otimes M\ar[dl, "\varepsilon\otimes\id_M"]\\
		k\otimes M
	\end{tikzcd}	
	\caption{Diagrams for the definition of a comodule.}
\end{figure}

\begin{rmk}
	Notationally speaking, we write $_V\mathcal{M}$ and $^V\mathcal{M}$ for the categories of 
	left modules and comodules. Similar for right boyes.
\end{rmk}

\subsection{Tensors of $V$-modules}
Say $V$ is a bialgebra (or Hopf if you prefer, but it's not necessary). Then we get the following:
\begin{lem}
	Let $M$ and $N$ be two right $V$-modules. Then
	\begin{enumerate}
		\item $M\otimes N$ is a right $V$-module.
		\item If $V$ is cocommutative, then $M\otimes N\cong N\otimes M$ as right $V$-modules.
		\item $\mathcal{M}_V$ is a tensor (monoidal) category.
	\end{enumerate}
\end{lem}
\begin{prf}
	For (a), use the fact that $V\otimes V$ acts on $M\otimes N$ in a natural way. Then you get
	a map $V\otimes V\to \Hom_k(M\otimes N,M\otimes N)^{op}$ (why opposite?) and by precomposing
	with $\Delta$ to show they are $V$-modules. This establishes (b).

	Finally (c) follows because $\Vectk$ is a tensor category.
\end{prf}

\subsection{Hopf Modules}\label{subsec-hopfmod}
\begin{defn}
	Let $V$ be a bialgebra over $k$. We say that $M$ is a $\binom{r}{r}$ Hopf $V$-module if
	\begin{enumerate}
		\item $(M,m_M)$ is a right $V$-module
		\item $(M,\rho_M)$ is a right $V$ comodule.
		\item $\rho_M$ is a right $V$-module map.
		\begin{itemize}
			\item Equivalently, $m_M$ is a right $V$-comodule map.
		\end{itemize}
	\end{enumerate}
\end{defn}
\begin{rmk}
	The fact that these two last conditions are equivalent are perhaps not immediately obvious but we
	have been assured it is true. :)
\end{rmk}

\section{January 16, 2019}
\begin{thm}[Larson-Sweedler `69]\label{thm-LS69}
	Let $V$ be a Hopf algebra over $k$. Then the following categories are equivalent:
	\[_V^V\mathcal{M}\cong \Vectk\cong \mathcal{M}_V^V\]
\end{thm}

\begin{rmk}
	A natural question that one may ask is how to extend this theorem to the enriched setting.
\end{rmk}

\begin{lem}
	$S$ is an anti-homomorphism of an algebra $V$ and an anti-homomorphism of the coalgebra $V$.
\end{lem}
\begin{prf}
	Use the facts proved in problem~\ref{prob-4.1}.
\end{prf}

\begin{prob}
	If $\ch k=0$ any Hopf quotient of $k[x]$ is either $k$ or $k[x]$ itself.
\end{prob}

\begin{prob}
	Suppose $k$ has positive characteristic. Construct two different Hopf algebra quotients of $k[x]$ of dimension $p=\ch k$.
\end{prob}

\begin{ex}
	$\ch k=p>0$. Let $V=k[x]/(x^p)$ as an algebra and define $\Delta(x)=x\otimes 1+1\otimes x$,
	$\varepsilon(x)=0$ and $S(x)=x$. Then notice
	\[\Delta(x^p)=\sum\binom{p}{i}x^i\otimes x^{p-i}=x^p\otimes 1+1\otimes x^p.\]
\end{ex}

\subsection{Hopf Modules Once More}
Recall the definitions from section~\ref{subsec-comod} and section~\ref{subsec-hopfmod}.

\begin{rmk}
	An equivalent (more category-theoretical) definition of a Hopf module is a $V$-comodule object in 
	the category of $V$-modules. Also can dualize everything.
\end{rmk}

\begin{ex}
	Consider $V\in _V^V\mathcal{V}$. THen you can define $\rho=\Delta$ and check it satisfies all the requirements.
\end{ex}
\begin{defn}
	An $\binom{l}{l}$ Hopf $V$-module $M$ is called \textbf{trivial} if $M\cong V\otimes M_0$ for some $M_0\in\Vectk.$
\end{defn}
\begin{rmk}
	Basically you can think of this being trivial since we can \textit{always} define a module in this way
	where the entire module structure is inherited from the structure of $V$ (that is, irrespective of $M_0$).
\end{rmk}
\begin{thm-prime}{thm-LS69}%[Fundamental Thm of Hopf Modules (L-S `69)]
	Suppose that $V$ is a Hopf algebra. Then
	\begin{enumerate}
		\item Every $\binom{l}{l}$ Hopf $V$-module is trivial.
		\item Let $M\in _V^V\mathcal{M}$. Then 
		\[M\xrightarrow{\sim} V\otimes M^{Cov}\]
		where
		\[M^{Cov}:=\{m\in M|\rho(m)=1\otimes m\}.\]
		\item $_V^V\mathcal{M}\cong \Vectk$.
	\end{enumerate}
\end{thm-prime}
\begin{rmk}
	This is actually just a reformulation of theorem~\ref{thm-LS69} in less compact (but ultimately more usable) notation.
\end{rmk}
\begin{prf}
	Some identites that will be useful:
	\begin{align*}
		\Delta(v)=\sum v_1\otimes v_2\\
		\sum S(v_1)\otimes v_2=\varepsilon(v)=\sum v_1S(v_2)\\
		\rho(m)=\sum m_{-1}\otimes m_1\\
		\sum v_1\otimes (v_2)_1\otimes (v_2)_2=\sum (v_1)_1\otimes (v_1)_2\otimes v_2=\sum v_1\otimes v_2\otimes v_3\\
		\sum (m_{-1})_1\otimes (m_{-1})_2\otimes m_0=\sum m_{-1}\otimes (m_0)_{-1}\otimes (m_0)_0=\sum m_{-2}\otimes m_{-1}\otimes m_0
	\end{align*}

	Then we defin $\phi:M\to M^{Cov}$ by 
	\[\phi(x)=\sum S(x_{-1})x_0\in M\]
	We claim first that $\phi(x)\in M^{Cov}$ -- namely $\rho(\phi(x))=1\otimes\phi(x)$. To see this, compute
	\begin{align*}
		\rho(\phi(x)) &= \sum (\phi(x))_{-1}\otimes (\phi(x))_0\\
		&= (S(x_{-1})x_0)_{-1}\otimes (S(x_{-1})x_0)\\
		&= \sum\Delta(S(x_{-1}))\rho(x_0)\\
		&= \sum (S\otimes S)\circ \tau(\Delta(x_{-2}))\cdot [x_{-1}\otimes x_0]\\
		&=\sum[S(x_{-2})\otimes S(x_{-3})][x_{-1}\otimes x_0]\\
		&= \text{my fingers are aching...}\\
		&= 1\otimes\phi(x)
	\end{align*}

	Now define $F:M\to V\otimes M^{Cov}$ by $F=(\id\otimes\phi)\circ\rho$ and $G:V\otimes M^{Cov}$ 
	to be the map taking $v\otimes m$ to $vm$. The next claim is that $GF$ and $FG$ are the identity.
	You can see this by similarly pushing around notation.

	Finally the last claim will be seen on Friday.
\end{prf}
\end{document}

### 2019-01-18 Lie Algebra Notes/Lie Algebras.tex
Update Notes

\begin{hwprob}{2.1}
	Prove that the set of all inner derivations $\ad x$ for $x\in L$ is an ideal of $\Der L$.
\end{hwprob}
\begin{sol}

\begin{hwprob}{3.8}
	Let $L$ be nilpotent. Prove that $L$ has an ideal of codimension 1.
\end{hwprob}
\begin{sol}

@@

\begin{defn}
	If $L^{(n)}$ for some $n$, and thus for all sufficiently large $n$, we call $L$ \textbf{solvable.}
\end{defn}
\begin{rmk}
	Note that actually there is a cooresponding definition for groups using commutators $[G,G]$.
	This was defined first and ported to Lie algebras. It actually happened in the reverse direction for:
\end{rmk}

@@

Using this definition, we are able to give two different characterizations:
\begin{defn}[Semisimple Lie Algebra]\label{def-semisimple}
	$L$ is semisimple if $\Rad L=0$.
\end{defn}
\begin{defprime}{def-semisimple}[Semisimple Lie Algebra]
	$L$ is semisimple if it is a direct sum of simple Lie algebras.
\end{defprime}
This equivalence will not become clear until we learn about the Killing form, but at least it jives
with our experience with representation theory and semisimple modules.
\end{document}

>>>>

\begin{hwprob}{2.1}
	Prove that the set of all inner derivations $\ad x$ for $x\in L$ is an ideal of $\Der L$.
\end{hwprob}
\begin{sol}
	Let $I=\ad L$. $I\subseteq \Der L$ since $\ad x$ is a derivation. This is a subspace since $\ad 0=0$ and since the bracket is bilinear
	\[\ad x(ay+bz)=[x,ay+bz]=[x,ay]+[x,bz]=a[xy]+b[xz]=a\ad x(y)+b\ad x(z).\]

	Take any $x\in I$ and let $f\in\Der L$ be arbitrary. But then for $y,z\in L$
	\begin{align*}
		[\ad x,f](yz)&=(\ad x f-f\ad x)(yz)\\
		&=[x,f(yz)]-f[x,yz]\\
		&=[x,f(y)z+yf(z)]-f([x,y]z+y[x,z])\\
		&=[x,f(y)z]+[x,yf(z)]-f([x,y]z)-f(y[x,z])\\
		&=[x,f(y)]z+f(y)[x,z]+[xy]f(z)+y[x,f(z)]-f([x,y])z\\
		&\hspace{2in} -[x,y]f(z)-f(y)[x,z]-yf[x,z]\\
		&=([x,f(y)]-f[x,y])z+y([x,f(z)]-f[x,z])\\
		&=([\ad x, f](y))z+y([\ad x,f](z))
	\end{align*}
	so $[IL]\subseteq I$ whence $I$ is an ideal of $L$.
\end{sol}

\begin{hwprob}{3.8}
	Let $L$ be nilpotent. Prove that $L$ has an ideal of codimension 1.
\end{hwprob}
\begin{sol}
	Assume that $L$ is nilpotent. Then defining $L^0=L$, and by induction $L^i=[LL^{i-1}]$, this 
	means that $L^n=0$.
\end{sol}

@@

\begin{defn}
	If $L^{(n)}=0$ for some $n$, and thus for all sufficiently large $n$, we call $L$ \textbf{solvable.}
\end{defn}
\begin{rmk}
	Note that actually there is a cooresponding definition for groups using commutators $[G,G]$.
	This was defined first and ported to Lie algebras. It actually happened in the reverse direction for:
\end{rmk}

@@

Using this definition, we are able to give two different characterizations:
\begin{defn}[Semisimple Lie Algebra]\label{def-semisimple}
	$L$ is semisimple if $\Rad L=0$.
\end{defn}
\begin{defprime}{def-semisimple}[Semisimple Lie Algebra]
	$L$ is semisimple if it is a direct sum of simple Lie algebras.
\end{defprime}
This equivalence will not become clear until we learn about the Killing form, but at least it jives
with our experience with representation theory and semisimple modules.

\section{January 14, 2019}
I missed this day for an event. I will try to catch up if I have time.

\section{January 16, 2019}
We are continuing from last time when we were showing that as solvable Lie subalgebra of $\gl(n,k)$
(when $\ch k=0$ and $\bar k=k$) necessarily admits a vector $v\in k^n\setminus 0$ that is
a common eigenvector:
\[xv=\lambda(x)v\]
for some \textbf{linear} function $\lambda:L\to k$.

We argues (as in the parllel case of a Lie subalgebra of $\gl(n,k)$ consisting of nilpotent matrices
by induction on dimension and found the same result (except it is zero). In this case when $\dim L=1$,
since $k$ is algebraically closed, $L$ is spanned by asingle matrix $x$ and any eigenvector of $x$ does the trick.

Now assume that the result holds for solvable Lie algebras of dimension $<d$. Then since $L$ is sovable, we know
$[LL]\subsetneq L$. Let $I$ be any subspace containing $[LL]$ and properly contained in $L$ of codimension one.
Then $I$ is an ideal since $[LI]\subseteq[LL]\subseteq I$, so by the inductive hypothesis $I$ has a common eigenvector, 
so that there is a linear function $\lambda$ such that the \textbf{weight space}
\[V_\lambda=\{v\in k^n:xv=\lambda(x)v,\,\forall x\in I\}\]
is nonzero.

We know that $L=I+kx$ for some $x\in L$, so we must show that $x$ preserves $V_\lambda$: then any eigenvector
of $x$ on $V_\lambda$ will do the job. Given $v\in V_\lambda$, we must show that 
\[yxv=xyv+[yx]v=\lambda(y)xv+\lambda([yx])v\]
so we must show that $\lambda([xy])=0$.

Given $v\in V_\lambda$, look at hte powers $v,xv, x^2v,\dots$. Let $m$ be the least power such that the
$v,xv,\dots,x^mv$ are linearly independent and let $W_i$ be the span of the first $i$ of these.

Then one checks by induction for an $y\in I$ that $yx^iv=\lambda(y)x^iv\pmod{v,xv,\dots,x^iv}$
It follows that $y$ was in the subspace $W_m$ (oh no missed some stuff).

So the trave of $y$ on $W_m$ is $m\lambda(y)$. Now for any $y\in I$, the matrix $[xy]$ acts on $W_m$ ascendingthe commutator of two matrices having trace zero, whence $m\lambda([xy])=0$,
so since $\ch k=0$, $\lambda([xy])=0$, as desired.

As a consequence, given a solvable Lie subalgebra of $\gl(n,k)$ here is a chain of subspaces
\[0=V_0\le V_1\le\cdots\le V_n=k^n\]
or a \textbf{flag} such that $LV_i\subseteq V_i$ for all $i$: $L$ acts by upper triangular 
matrices with respect to a suitable basis.

Another way of putting this, using module language, is that given any \textbf{finite dimensional} module over
a solvable Lie algebra (over an algebraically closed characteristic zero field) has a one-dimensional submodules $N$.

Equivalently, the only finite-dimensional irreducible modules over a solvable Lie algebra are one-dimensional. We 
will use this later in the context of semisimple Lie algebras with large solvable subalgebra.

\begin{rmk}
	For the record, Lie's theorem \textit{definitely fails} when $k$ has positive characteristic.
	We see one in our homework, and another example comes from $\sl(2,k)$ where $\ch k=2.$

	We know that $\sl(2,k)$ has a basis: $(\begin{smallmatrix}
		1&0\\0&-1
	\end{smallmatrix})=I_2$, $x=e_{12}$, $y=e_{21}$. Here $[hx]=0=[hy]$ and $[xy]=h$. It is shown in the book
	and is easy to see that when $\ch k\ne 2$, $\sl(2,k)$ is simple, but here $\sl(2,k)$ is nilpotent and its
	bracket relations become $[xy]=z$, $[xz]=[yz]=0$.

	It ends up we can define (actually over any field) the \textbf{Heisenberg} Lie algebra, spanned
	by $x=e_{12},y=e_{23},z=e_{13}$ or more simply the $3\times 3$ strictly-upper-triangular matrices.

	The matrices in $\sl(2,k)$ for any $k$ have no common eigenvector, so our proof of Lie's theorem must
	fail somewhere? We indeed have an ideal $I$ of coimension one spanned by $h$ and $x$ and there is acommon eigenvector for $I$, 
	namely $e_1$. Then $I$ acts diagonally on $ke_1$ via the linear function sending $h$ to 1, $x$ to $1$. 
	Here $y$ plays the role of $x$ in the proof, but is outside the ideal. That is,
	\[\lambda[yx]=\lambda(h)\ne 0.\]
\end{rmk}

Returning to the good case of $k$, we can now apply our result on a linear solvable Lie algebras to arbitrary ones:
uGIven any solvable Lie algebra $L$, there is a flag $0=L_0\le L_1\le\cdots\le L_n=L$ of ideas of $L$
so that $[LL_i]\subseteq L_i$ Also $[LL]$ is nilpotent, since it acts on $L$ by strictly upper triangular matrices.

Both results fail for general solvable Lie algebras when the basefield is not nice. If a solvable Lie algebra $L$ 
\textbf{does} have a chain of ideals $0=L_0\le L_1\le\cdots\le L_n=L$ where $\dim L_i=i$, then we call $L$ \textbf{completely solvable.}

\subsection{Returning to Linear Algebra}
At this point we will take some time to return to an important idea from Linear algebra, namely the \textbf{Jordan canonical form}
of a square matrix. Recall Jordan blocks and similarity when eigenvalues lie in the field.

If $M$ is already in this form, then we can write $M=M_s+M_n$ where $M_s$ consists of the diagonal entries of each
block (zeroes elsewhere) and $M_n$ the off-diagonal entries. Then $M_s$ is diagonal (or for general $M$
diagonalizable), or \textbf{semisimple} and $M_n$ is nilpotent and $[M_sM_n]=0$ (they commute).

\begin{defn}
	$M=M_s+M_n$ is called the \textbf{Jordan decomposition of $M$.} We call $M_s$ and $M_n$ the \textbf{semisimple} 
	and \textbf{nilpotent} parts of $M$.
\end{defn}
\begin{rmk}
	Each are uniquely determined for $M$ and (and this is not generally proved in the first year 
	algebra sequence) each are equal to polynomials in $M$ with no constant term.x`
\end{rmk}

\end{document}

### 2019-01-18 Reading Notes/reading_notes.tex
Update Notes

>>>>

Consider the representations of A over a field $k$ (or more generally $A$ modules. W can make this a tensor catecory)
Recall that in this category we have enough projectives and (with finitely generated hopf algebras) we get that all projectives are injective.

Then we end up defining $\Ext_A^*(k,k)=H^*_A(A,k)$, the cohomology ring of $A$ over $k$.

The conjecture here is that when $A$ is finite dimensional that $H^*(A,k)$ is finitely generated as a $k$-algebra.

This is known in some special cases:
\begin{itemize}
	\item Finite groups
	\item Finite group schemes (in positive characteristic) due to Friedlander and Suslin.
	\item In characteristic zero we have
	\begin{itemize}
		\item Quantum groups
		\item Hopf algebras that come from Nichols algebras
	\end{itemize}
\end{itemize}

Papers to look at: Ginzburg and Kumar '94 Something like ``Cohomology of small quantum groups''. There is also a '10 paper on Julia's website with Sarah Witherspoon and two others on Nichols algebras.

Sarah Witherspoon has a book she is writing on Hochschild cohomology that could be very good. It's on her website (Texas A\& M). Check out chapter 9 and appendix A.
